---
title: "p8105_hw2_bes2161"
author: "Brittany Shea"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)
```

### Problem 1

```{r}
trans_ent = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The dataset shows NYC Transit data with information on entrances and exits for subway stations. These data are not tidy because route and route # should be variables. We should convert `route` variables from wide to long format. So far we have imported data, updated variable names, and selected the columns to use. We updated `entry` from `yes` / `no` to a logical variable. We specify that `Route` columns 8-11 should be character. There are 20 columns and 1868 rows.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```
There are 465 unique stations.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
There are 84 ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
The proportion of station entrances / exits without vending that allow entrance is 0.377.

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
60 stations serve the A train, and 17 are ADA compliant.
